---
title: "Working with Databases - Part II"
author: "Dr. Hua Zhou @ UCLA"
date: "Feb 18, 2021"
subtitle: Biostat 203B
output:
  html_document:
    toc: true
    #toc_depth: 2
  #ioslides_presentation:
  #  smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 3.5, fig.align = 'center',
  cache = FALSE, cache.lazy = FALSE)
```

## Introduction

In this lecture we will learn:

- Import data from bigish csv files (MIMIC-III).

- Tidy data (TODO). 

- Deposit data into an SQLite database.

- Query SQLite database. 

- Transform in database and plot in R.

## Machine information

```{r}
sessionInfo()
```
```{r}
library(tidyverse)
```

## CSV file

The `/usr/203b-data/mimic-iii` folder on the teaching server contains the MIMIC-III data.
```{bash, eval = (Sys.info()[["sysname"]] == "Linux")}
# On teaching server
ls -l /home/203bdata/mimic-iii
```
On my Mac, it's at `/Users/huazhou/Google Drive/mimic/mimic-iii`.
```{bash, eval = (Sys.info()[["sysname"]] == "Darwin")}
# On my MacBook
ls -l /Users/huazhou/Google\ Drive/mimic/mimic-iii
```
```{r}
datafolder <- ifelse(Sys.info()[["sysname"]] == "Linux",
                    "/usr/203b-data/mimic-iii",
                    "/Users/huazhou/Google\ Drive/mimic/mimic-iii")
list.files(datafolder, full.names = TRUE, patter = "*.csv")
```

## Read csv file: `read.csv()` vs `read_csv()` (tidyverse) vs `fread()` (data.table)

In HW2, we learnt that the `fread()` function in the data.table package is much more efficient for parsing large text files than `read.csv()` in base R.

## Read all CSVs and deposit to an SQLite database

Create a SQLite database file called `mimiciii.sqlite`.
```{r}
library("DBI")
library("RSQLite")
con = dbConnect(RSQLite::SQLite(), "mimiciii.sqlite")
str(con)
```


Because of excellent efficiency of `data.table` packages, we use it to read all CSVs and deposit into `mimiciii.sqlite`. **Do not run this chunk by yourself on teaching server.** This takes a couple minutes.
```{r, eval = (!file.exists("mimiciii.sqlite") | file.size("mimiciii.sqlite") == 0) & (Sys.info()[["sysname"]] != "Linux")}
datafolder
csv_names <- list.files(datafolder)
csv_names <- csv_names[str_detect(csv_names, "csv")]
csv_names
system.time({
for (table_name in csv_names) {
    csv_name <- str_c(datafolder, "/", table_name)
    tmp <- fread(csv_name)
    table_name <- str_remove(table_name, ".csv")
    dbWriteTable(con, table_name, tmp, overwrite = TRUE)
}})
```
```{r}
# list tables in database
dbListTables(con)
# disconnect from database
dbDisconnect(con)
```
The resultant database file `mimiciii.sqlite` has size 10.4GB.
```{bash}
ls -l mimiciii.sqlite
```
On the teaching server, this file is located at `/usr/203b-data/mimic-iii-db/`.

## Read data from database

Connect to the database `mimiciii.sqlite` and list the tables:
```{r}
con <- dbConnect(RSQLite::SQLite(), 
                 dbname = ifelse(Sys.info()[["sysname"]] == "Linux",
                    "/usr/203b-data/mimic-iii-db/mimiciii.sqlite",
                    "./mimiciii.sqlite"))
dbListTables(con)
```

Read the table `CHARTEVENTS`:
```{r}
chartevents <- tbl(con, "CHARTEVENTS")
class(chartevents)
chartevents %>% print(width = Inf)
```

## Use dplyr with SQLite

Retrieve chart events of `SUBJECT_ID` beetween 23 and 164:
```{r}
chartevents_subset <- chartevents %>%
  select(ROW_ID, SUBJECT_ID, HADM_ID, ITEMID, ITEMID) %>%
  filter(SUBJECT_ID >= 23, SUBJECT_ID <= 165)
chartevents_subset
```

## SQL query

```{r}
class(chartevents_subset)
show_query(chartevents_subset)
```

## SQL grouping

```{r}
chartevents %>%
  group_by(SUBJECT_ID) %>%
  summarise(n = n())
```

## SQL query

```{r}
chartevents %>%
  group_by(SUBJECT_ID) %>%
  summarise(n = n()) %>%
  show_query()
```

## SQL translation

dbplyr package (a dplyr backend for databases) has a function, `translate_sql`, that lets you experiment with how R functions are translated to SQL:
```{r}
library("dbplyr")
translate_sql(x == 1 & (y < 2 | z > 3))
translate_sql(x ^ 2 < 10)
translate_sql(x %% 2 == 10)
translate_sql(paste(x, y))
translate_sql(mean(x))
translate_sql(mean(x, na.rm = TRUE))
```

## Timings

Let's compare the timings of dplyr (in-momory) and dbplyr (on disk database).

- dplyr using tibble:
```{r}
csvfile <- str_c(datafolder, "/DATETIMEEVENTS.csv")
dtevents_tibble <- read_csv(csvfile)
system.time(
  dtevents_tibble %>%
  select(ROW_ID, SUBJECT_ID, HADM_ID, ITEMID) %>%
  # filter(SUBJECT_ID >= 23, SUBJECT_ID <= 165) %>%
  group_by(SUBJECT_ID) %>%
  summarize(n = n())
)
```
- dplyr using SQLite:
```{r}
dtevents_sql <- tbl(con, "DATETIMEEVENTS")
system.time(
  dtevents_sql %>%
  select(ROW_ID, SUBJECT_ID, HADM_ID, ITEMID) %>%
  # filter(SUBJECT_ID >= 23, SUBJECT_ID <= 165) %>%
  group_by(SUBJECT_ID) %>%
  summarize(n = n())
)
```

SQLite (0.08 seconds) was much faster than tibble (0.25 seconds). But SQLite is disk-based, while the tibble is in memory. Why is the discrepancy?

## Laziness

dplyr/dbplyr uses lazy evaluation as much as possible, particularly when working with non-local backends.

- When building a query, we don’t want the entire table, often we want just enough to check if our query is working.

- Since we would prefer to run one complex query over many simple queries, laziness allows for verbs to be strung together.

- Therefore, by default dplyr

    - won’t connect and query the database until absolutely necessary (e.g. show output),

    - and unless explicitly told to, will only query a handful of rows to give a sense of what the result will look like
    
```{r}
dtevents_sql %>%
  select(ROW_ID, SUBJECT_ID, HADM_ID, ITEMID) %>%
  # filter(SUBJECT_ID >= 23, SUBJECT_ID <= 165) %>%
  group_by(SUBJECT_ID) %>%
  summarize(n = n())
```

## Full query

To force a full query and return a complete table it is necessary to use the `collect` function.
```{r}
system.time(
  dtevents_sql %>%
  select(ROW_ID, SUBJECT_ID, HADM_ID, ITEMID) %>%
  # filter(SUBJECT_ID >= 23, SUBJECT_ID <= 165) %>%
  group_by(SUBJECT_ID) %>%
  summarize(n = n()) %>%
  collect()
)
```

## Plotting

### Bar plot

Suppose we want the bar plot of number of datetime events on each `SUBJECT_ID`.
```{r}
chartevents %>%
  count(SUBJECT_ID) %>%
  ggplot() + 
  geom_histogram(mapping = aes(x = n))
```
Apparently current version of ggplot2 (v3.2.1) works for `tbl_dbi/tbl_sql/tbl_lazy` now. 

Older version of ggplot2, e.g., v2.2.1, will output error. This is becasue ggplot2 needed to compute the count per bin by going through all the rows. But here `chartevents` is just a pointer to the SQLite table. We had to use the **transform in database, plot in R** strategy.
```{r}
chartevents %>%
  count(SUBJECT_ID) %>%
  collect() %>%
  ggplot() + 
  geom_histogram(mapping = aes(x = n))
```

### Other plots

For example of making histogram and raster plot, read tutorial  <https://db.rstudio.com/best-practices/visualization/>.

## Close connection to database

```{r}
dbDisconnect(con)
```